Radial and rotational velocities of young brown dwarfs and very low-mass stars in the Upper Scorpius OB association and the $\rho$~Ophiuchi cloud core

### 2006-05-09 BD_radvel/kurosawa.tex
Change after Tim's revision #2.  RK
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@715 25464f21-6099-402c-b30a-640f1d40d38b

We present the results of a radial velocity (RV) survey of 14 brown
dwarfs (BDs) and very low-mass (VLM) stars in the Upper Scorpius OB
association (UScoOB) and 3 BD candidates in the $\rho$~Ophiuchi dark
cloud core. We obtained high-resolution echelle spectra at VLT using
Ultraviolet and Visual Echelle Spectrograph (UVES) at two different
epochs for each object, and measured the shifts in their RVs to
identify candidates for binary/multiple systems in the sample. The
average time separation of the RV measurements is 21.6~d, and our
survey is sensitive to the binaries with separation $<0.12$~au.  We
found 4 out of 17 objects (or $24^{+16}_{-14}$~per~cent by fraction)
show a significant RV change in 4--33~d time scale, and are 
considered as binary/multiple `candidates.'  We found no double-lined
spectroscopic binaries in our sample, based on the shape of
cross-corelation curves. The RV dispersion of the objects in UScoOB
is found to be very similar to that of the BD and VLM stars in
Chamaeleon~I (Cha~) previously studies by others. We also found the
distribution of the mean rotational velocities ($v\,\sin i$) of the
UScoOB objects is similar to that of the Cha~I, but the dispersion of
$v\,\sin i$ is significantly larger than that of the Cha~I objects.

@@

Next, we investigate the range of semimajor axes of binaries
(equivalently the range of binary periods) to which our RV survey is
sensitive. For this purpose, we will consider the detection
probability for binaries or the RV variables, given the time
separations of two-epoch observations and the ranges of estimated
primary masses (c.f.~Tables~\ref{tab:literatures} and
\ref{tab:summary_results}).  The probability will be calculated based
on the simulated RV observations of binaries whose orbit are randomly
selected from a model. A similar method was used by
\citet{maxted:2005} and \citet{basri:2006}.  The most important
factors in determining the detection probability are the size of
uncertainties in RV measurements (which we use the average
$\sigma_{\mathrm{RRV}}$ from our observation), and the time
separations of observations.  The smaller the errors in RV
measurements, the larger the probability for a given binary orbit and
a time separation of orbit. The larger the time separation of
observations, the larger the upper limit of the semimajor axis to
which an observation is sensitive.  The average time separation of the
two-epoch RV observations of our targets is $21.6$~d.
In the following, we will briefly discuss our model assumptions and
parameters which are essentially the same as those of
\citet{maxted:2005} but with some simplifications.

@@

In Section~\ref{sub:Radial-velocities}
(see Fig.~\ref{fig:chisq-test}), we found the histogram of the
$\chi^{2}$ probability $p$, for the RV data fit assuming constant RV,
reasonably matches the expected distribution.  This lead us to
conclude that the uncertainty estimates in RV values were
reasonable. Here, we will inspect the effect of the systematic error
on the observed binary fraction more closely. There are two possible
source of errors: (1)~the RV shifts caused by surface/atmospheric
activity, and (2)~the RV shifts due to the limit of the stability of
autoguiding of the telescope.  \citet{joergens:2006} found that (1) is
negligibly small for a young VLM objects with $M_{*}<0.12\,M_{\sun}$.
The size of RV errors caused by (2) is expected to be $<1\,\kmps$
\citep{bailer-jones:2004}. Note that \citet{maxted:2005} found an
additional systematic error was not needed to be added to the RV
uncertainties of \citet{joergens:2006} who used the same instruments
as ours, and used the same method of estimating the
uncertainties. Hence, we expect the systematic error to be much
smaller than $<1\,\kmps$ if one is required.

The observed binary fraction is expected to be sensitive to the
assumed size of the systematic error ($\sigma_{\mathrm{sys}}$) since
the RV shifts and the systematic errors are in the same order of
magnitude. The number of binary detections would decrease as
$\sigma_{\mathrm{sys}}$ increases.  We have computed the $\chi^2$
probabilities ($p$) for each object being a RV constant object as done
earlier (Section\ref{sub:Radial-velocities}), but with the assumed 
systematics error of $\sigma_{\mathrm{sys}}=0.2\,\kmps$, which was added to
the uncertainties in the RVs of each object.  The results are
summarised in Table~\ref{tab:prob_compare} along with the probabilities
computed without the systematic error (as in
Section\ref{sub:Radial-velocities}).  The number of objects with
$p<10^{-3}$ decreases from 4 to 2 as the systematic error is
introduced. Unfortunately, the probabilities computed with
$\sigma_{\mathrm{sys}}=0.0\,\kmps$ for GY~141 and USco~55 are just
below $p=10^{-3}$ boundary value; hence, they become RV constant
objects as $\sigma_{\mathrm{sys}}=0.2\,\kmps$ is introduced. This
clearly demonstrates the importance of the uncertainty estimates in
the measured RV values.  We found that the histogram of the $\chi^2$
probabilities with $\sigma_{\mathrm{sys}}=0.2\,\kmps$ (as in
Fig~\ref{fig:chisq-test}), also have a good agreement with the
expected distribution of $-\log{p}$ (not shown here).  
The binary fraction estimated with the assumed systematic error is
$2/17$ or $f=12^{+13}_{-10}$~per~cent for $a<0.12$~au.

@@

We have presented two-epoch RV survey of 18 young BDs and VLM stars
($0.02~M_{\sun} < M_{*} < 0.1~M_{\sun}$) in
UScoOB and $\rho$~Oph dark cloud core using the high resolution UVES
echelle spectroscopy at VLT.  The average time separation of RV
measurements are 21.6~d, and our RV survey is sensitive to binaries
with separation smaller than 0.12~au.  One of our targets USco~121 is
most likely a non-member of the UScoOB association based on the
deviation of the RV from the result of the population in the
association. A similar conclusion was found by \citet{muzerolle:2003}
from their RV study and the low lithium abundance.

We found 4 (GY~141, GY~310, USco~55 and USco~104) out of 17 objects as our 
binary/multiple candidates. This corresponds to the binary fraction of
$24^{+16}_{-14}$~per~cent for the binary separation $a<0.12$~au.  We found
that the estimated binary fraction is very sensitive to
the assumed size of systematic error. If an additional systematic
error of $0.2\,\kmps$  was added to the uncertainties in the RV
measurements, we found only 2 (GY~310 and USco~104) out of 17 objects
as our binary/multiple candidates, which corresponds to the binary
fraction of $12^{+13}_{-10}$~per~cent.

The recent high-resolution imaging survey of brown dwarfs and very
low-mass objects (M5.5--M7.5 ) in the UScoOB by \citet{kraus:2005} (which was
not known to authors at the time of our observation: April--May, 2004)
confirms that USco~55 and USco~66 are multiple systems, and USco~109
is most likely a multiple system.  Their projected separations are
($a>4.0$~au), and are well above our sensitivity limit of
0.12~au. Hence, it is not surprising that we did not find USco~66 and
USco~109 as multiples. Interestingly, we also found USco~55 as a
candidate for a multiple system. This may indicate that USco~55 is  
a triple system with one of object located within 0.12~au from the
primary.  In addition, we found USco~104 as a binary candidate, but this
was not in their target list.  Further they found USco~67, 75, 112, 128,
130, 132 as non-multiple, and similarly we found them as
'non-candidates' for multiplicity.

We found the RV dispersion ($1.2\,\kmps$) of the objects in UScoOB is
very similar to that of the BDs and VLM stars in Chamaeleon~I (Cha~I)
previous study by \citet{joergens:2006b}. The rotational velocities
($v\,\sin i$) of the samples are also measured. The distribution of
$v\,\sin i$ for the UScoOB objects peaks around $16.9\,\kmps$ which is
also similar to that of the Cha~I population found by
\citet{joergens:2006b}; however, the dispersion of $v\,\sin i$ for the
UScoOB objects ($27.8\,\kmps$) is found to be much larger than that of
the Cha~I objects ($8.0\,\kmps$)

The follow-up spectroscopic observations of the binary candidates
presented here are planned in near future. There are only a few RV
variable binary candidates are identified from the earlier surveys
(\citealt{guenther:2003}; \citealt{kenyon:2005}; \citealt{joergens:2006}).
As \citet{burgasser:2006} pointed out most of the current RV and
imaging surveys use samples from magnitude-limited survey, but one
should attempt to use the samples from volume-limited survey in order
to find a correct statistics on binary parameters more straight forwardly,
i.e. without correcting for bias.

@@

We thank the staff of VLT of the ESO for the observations carried out in
service mode.  This work is supported by PPARC rolling grant PP/C501609/1.

>>>>

We present the results of a radial velocity (RV) survey of 14 brown
dwarfs (BDs) and very low-mass (VLM) stars in the Upper Scorpius OB
association (UScoOB) and 3 BD candidates in the $\rho$~Ophiuchi dark
cloud core. We obtained high-resolution echelle spectra at VLT using
Ultraviolet and Visual Echelle Spectrograph (UVES) at two different
epochs for each object, and measured the shifts in their RVs to
identify candidates for binary/multiple systems in the sample. The
average time separation of the RV measurements is 21.6~d, and our
survey is sensitive to the binaries with separation $<0.12$~au.  We
found 4 out of 17 objects (or $24^{+16}_{-14}$~per~cent by fraction)
show a significant RV change in 4--33~d time scale, and are 
considered as binary/multiple `candidates.'  We found no double-lined
spectroscopic binaries in our sample, based on the shape of
cross-correlation curves. The RV dispersion of the objects in UScoOB
is found to be very similar to that of the BD and VLM stars in
Chamaeleon~I (Cha~) previously studies by others. We also found the
distribution of the mean rotational velocities ($v\,\sin i$) of the
UScoOB objects is similar to that of the Cha~I, but the dispersion of
$v\,\sin i$ is significantly larger than that of the Cha~I objects.

@@

Next, we investigate the range of semimajor axes of binaries
(equivalently the range of binary periods) to which our RV survey is
sensitive. For this purpose, we will consider the detection
probability for binaries or the RV variables, given the time
separations of two-epoch observations and the ranges of estimated
primary masses (c.f.~Tables~\ref{tab:literatures} and
\ref{tab:summary_results}).  The probability is calculated based
on the simulated RV observations of binaries whose orbit are randomly
selected from a model. A similar method was used by
\citet{maxted:2005} and \citet{basri:2006}.  The most important
factors in determining the detection probability are the size of
uncertainties in RV measurements (which we use the average
$\sigma_{\mathrm{RRV}}$ from our observation), and the time
separations of observations.  The smaller the errors in RV
measurements, the larger the probability for a given binary orbit and
a time separation of orbit. The larger the time separation of
observations, the larger the upper limit of the semimajor axis to
which an observation is sensitive.  The average time separation of the
two-epoch RV observations of our targets is $21.6$~d.
In the following, we will briefly discuss our model assumptions and
parameters which are essentially the same as those of
\citet{maxted:2005} but with some simplifications.

@@

In Section~\ref{sub:Radial-velocities}
(see Fig.~\ref{fig:chisq-test}), we found the histogram of the
$\chi^{2}$ probability $p$, for the RV data fit assuming constant RV,
reasonably matches the expected distribution.  This leads us to
conclude that the uncertainty estimates in RV values were
reasonable. Here, we will inspect the effect of the systematic error
on the observed binary fraction more closely. There are two possible
source of errors: (1)~the RV shifts caused by surface/atmospheric
activity, and (2)~the RV shifts due to the limit of the stability of
autoguiding of the telescope.  \citet{joergens:2006} found that (1) is
negligibly small for a young VLM objects with $M_{*}<0.12\,M_{\sun}$.
The size of RV errors caused by (2) is expected to be $<1\,\kmps$
\citep{bailer-jones:2004}. Note that \citet{maxted:2005} found an
additional systematic error was not needed to be added to the RV
uncertainties of \citet{joergens:2006} who used the same instruments
as ours, and used the same method of estimating the
uncertainties. Hence, we expect the systematic error to be much
smaller than $<1\,\kmps$ if one is required.

The observed binary fraction is expected to be sensitive to the
assumed size of the systematic error ($\sigma_{\mathrm{sys}}$) since
the RV shifts and the systematic errors are in the same order of
magnitude. The number of binary detections would decrease as
$\sigma_{\mathrm{sys}}$ increases.  We have computed the $\chi^2$
probabilities ($p$) for each object being a RV constant object as done
earlier (Section\ref{sub:Radial-velocities}), but with the assumed 
systematics error of $\sigma_{\mathrm{sys}}=0.2\,\kmps$, which was added to
the uncertainties in the RVs of each object.  The results are
summarised in Table~\ref{tab:prob_compare} along with the probabilities
computed without the systematic error (as in
Section\ref{sub:Radial-velocities}).  The number of objects with
$p<10^{-3}$ decreases from 4 to 2 as the systematic error is
introduced. Unfortunately, the probabilities computed with
$\sigma_{\mathrm{sys}}=0.0\,\kmps$ for GY~141 and USco~55 are just
below the $p=10^{-3}$ boundary value; hence, they become RV constant
objects as the systematic error is introduced. This
clearly demonstrates the importance of the uncertainty estimates in
the measured RV values.  We found that the histogram of the $\chi^2$
probabilities with $\sigma_{\mathrm{sys}}=0.2\,\kmps$ (as in
Fig~\ref{fig:chisq-test}), is also in reasonable agreement with the
expected distribution of $-\log{p}$ (not shown here).  
The binary fraction estimated with the assumed systematic error is
$2/17$ or $f=12^{+13}_{-10}$~per~cent for $a<0.12$~au.

@@

We have presented two-epoch RV survey of 18 young BDs and VLM stars
($0.02~M_{\sun} < M_{*} < 0.1~M_{\sun}$) in
UScoOB and $\rho$~Oph dark cloud core using the high resolution UVES
echelle spectroscopy at VLT.  The average time separation of RV
measurements are 21.6~d, and our RV survey is sensitive to binaries
with separation smaller than 0.12~au.  One of our targets, USco~121, is
most likely a non-member of the UScoOB association based on the
deviation of the RV from the result of the population in the
association. A similar conclusion was found by \citet{muzerolle:2003}
from their RV study and the low lithium abundance.

We found 4 (GY~141, GY~310, USco~55 and USco~104) out of 17 objects as our 
binary/multiple candidates. This corresponds to the binary fraction of
$24^{+16}_{-14}$~per~cent for the binary separation $a<0.12$~au.  We found
that the estimated binary fraction is very sensitive to
the assumed size of systematic error. If an additional systematic
error of $0.2\,\kmps$  is added to the uncertainties in the RV
measurements, we find only 2 (GY~310 and USco~104) out of 17 objects
as our binary/multiple candidates, which corresponds to the binary
fraction of $12^{+13}_{-10}$~per~cent.

The recent high-resolution imaging survey of brown dwarfs and very
low-mass objects (M5.5--M7.5 ) in the UScoOB by \citet{kraus:2005} (which was
not known to authors at the time of our observation: April--May, 2004)
confirms that USco~55 and USco~66 are multiple systems, and USco~109
is most likely a multiple system.  Their projected separations are
$a>4.0$~au which is well above our sensitivity limit of
0.12~au; therefore, it is not surprising that we did not identify USco~66 and
USco~109 as multiples. Interestingly, we also found USco~55 as a
candidate for a multiple system. This may indicate that USco~55 is  
a triple system with one of object located within 0.12~au from the
primary.  In addition, we identify USco~104 as a binary candidate, but this
was not in their target list.  Further they found USco~67, 75, 112, 128,
130, 132 as non-multiple, and similarly we found them as
'non-candidates' for multiplicity.

We found the RV dispersion ($1.2\,\kmps$) of the objects in UScoOB is
very similar to that of the BDs and VLM stars in Chamaeleon~I (Cha~I)
previous study by \citet{joergens:2006b}. The rotational velocities
($v\,\sin i$) of the samples were also measured. The distribution of
$v\,\sin i$ for the UScoOB objects peaks around $16.9\,\kmps$ which is
also similar to that of the Cha~I population found by
\citet{joergens:2006b}; however, the dispersion of $v\,\sin i$ for the
UScoOB objects ($27.8\,\kmps$) is found to be much larger than that of
the Cha~I objects ($8.0\,\kmps$)

Follow-up spectroscopic observations of the binary candidates
presented here are planned in near future. There are only a few RV
variable binary candidates identified in earlier surveys
(\citealt{guenther:2003}; \citealt{kenyon:2005}; \citealt{joergens:2006}).
As \citet{burgasser:2006} points out most of the current RV and
imaging surveys use samples from magnitude-limited survey, but one
should attempt to use the samples from volume-limited survey in order
to find a correct statistics on binary parameters more straightforwardly,
i.e. without correcting for bias.

@@

We thank the staff of VLT of the ESO for carrying out the observations
in service mode. RK thank Rob Jeffries for helpful suggestions on the
data analysis presented in this paper. This work was supported by
PPARC rolling grant PP/C501609/1.
